MA_Results

### 2022-07-20 Markdown/Manuscript/Child_RMDs/MS_Intro_Short.Rmd
Dave and Kate's edits

<!-- `r ifelse(is.na(params$X@background_1L_chr),"",params$X@background_1L_chr)`  -->
Health economic computational models have become essential tools for health policy development [@dakin2015influence; @Erdemir2020]. Although influential and widely used, these models routinely contain errors [@Radeva2020], are rarely adequately validated [@Ghabri2019], can be difficult to reproduce [@Jalali2021; @McManus2019; @Bermejo2017] and are likely to be infrequently updated or revised [@Sampson2017]. To help address these issues, there is growing support for greater use of open source health economic models (OSHEMs) that grant liberal permissions to access and re-use model source code and data [@Pouwels2022]. However, to date actual implementations of OSHEMs are rare [@Emerson2019; @Michalczyk2018; @Feenstra2022]. Barriers to adoption include concerns about intellectual property, confidentiality, model misuse and the resources required to support open source implementations [@Pouwels2022; @Wu2019]. As many health economic models are owned by pharmaceutical companies and consultancies, commercial considerations may also limit the uptake of OSHEMs [@Feenstra2022].

Better guidance about how to implement OSHEMs is required [@Sampson2019]. Adherence to explicit guidelines is as essential requirement for modelling in healthcare [@Erdemir2020], but current best practice recommendations for OSHEMs are scarce and piecemeal. Guidelines on health economic model transparency were published ten years ago [@Eddy2012] and made recommendations on documenting models; however, notably missing were recommendations on the sharing of model code and data. More recent and more general modelling guidance [@Erdemir2020] recommends the sharing of code and data through platforms such as GitHub [@github2007] and Zenodo [@Zenodo2013] and the use of version control systems such as Git [@git20XX]. A coding framework for OSHEMs developed in the language R includes standardised approaches to directory structure and naming conventions [@Alarid2019].

>>>>

<!-- `r ifelse(is.na(params$X@background_1L_chr),"",params$X@background_1L_chr)`  -->
Health economic computational models have become essential tools for health policy development [@dakin2015influence; @Erdemir2020]. Although influential and widely used, these models routinely contain errors [@Radeva2020], are rarely adequately validated [@Ghabri2019], can be difficult to reproduce [@Jalali2021; @McManus2019; @Bermejo2017] and are likely to be infrequently updated or revised [@Sampson2017]. To help address these issues, there is growing support for greater use of open source health economic models (OSHEMs) that grant liberal permissions to access and re-use model source code and data [@Pouwels2022]. However, to date, actual implementations of OSHEMs are rare [@Emerson2019; @Michalczyk2018; @Feenstra2022]. Barriers to adoption include concerns about intellectual property, confidentiality, model misuse and the resources required to support open source implementations [@Pouwels2022; @Wu2019]. As many health economic models are owned by pharmaceutical companies and consultancies, commercial considerations may also limit the uptake of OSHEMs [@Feenstra2022].

Better guidance about how to implement OSHEMs is required [@Sampson2019]. Adherence to explicit guidelines is an essential requirement for modelling in healthcare [@Erdemir2020], but current best practice recommendations for OSHEMs are scarce and piecemeal. Guidelines on health economic model transparency were published ten years ago [@Eddy2012] and made recommendations on documenting models; however, notably missing were recommendations on the sharing of model code and data. More recent and more general modelling guidance [@Erdemir2020] recommends the sharing of code and data through digital repository services such as GitHub [@github2007] and Zenodo [@Zenodo2013] and the use of version control systems such as Git [@git20XX]. A coding framework for OSHEMs developed in the language R includes standardised approaches to directory structure and naming conventions [@Alarid2019].

### 2022-07-20 Markdown/Manuscript/Child_RMDs/MS_Results.Rmd
Dave and Kate's edits

Mental disorders impose high health, social and economic burdens worldwide [@RN8;@GBD2019]. Much of this burden is potentially avertable [@RN25], but poorly financed and organised mental health systems are ill-equipped for this challenge[@RN22;@RN23]. The large and widespread additional mental health burdens recently observed during the COVID-19 pandemic[@20211700] and predicted as a potential future consequence of global heating [@page_howard_2010], highlight the need to improve the resilience and adaptability of these systems. To help stem growing demand for mental health services, policymakers have also been encouraged to place greater emphasis on tackling the social determinants of mental disorder[@RN11].

<!-- Computational modelling could play an important role in developing policies to improve population mental health but this may require significant changes in the way mental health modelling projects are funded, conceptualised and implemented.  -->
Open source frameworks have been recommended for the development of mental health modelling field [@RN73] but, as with health economics more generally, OSHEMs remain rare. Currently there is only one mental health related model (in Alcohol Use Disorder [@Basu2018]) that is indexed in the Open Source Models Clearinghouse [@OSMC_20xx;@Emerson2019]. A Major Depressive Disorder reference model for the United States [@IVIMDD2022] is being developed as part of the Open Value Initiative [@Jansen2019]. Greater use of open source approaches could help improve the scope, validity and usefulness of mental health economic models.

Major mental health reform programs, such as those currently being implemented in Victoria, Australia [@VicRC2021], can involve the identification, prioritisation, sequencing, targeting and monitoring of multiple interdependent initiatives. Single purpose models that assume static systems may be inadequate for the decision support needs of policymakers and service planners [@PC2020]. Mental health economic models predominantly address issues relating to the affordability and value for money of individual programs [@RN34] with mental health simulation studies rarely modelling complex dynamic systems [@RN73]. Systems modelling approaches, that have been recommended for public health economics more generally [@SQUIRES2016588] and for mental health specifically [@Occhipinti2021], could provide greater insight about inter-dependencies between candidate policies and the dynamic nature of the mental health systems planning context. Dynamic systems methods might provide the foundations for developing reference models [@Afzali2013] of mental health systems that are intended for multiple-applications and re-use by different modelling teams. However, more complex models may be more prone to error [@Saltelli2019] and models designed for multiple purposes require greater investments in model transparency and validation [@Eddy2012; @Feenstra2022].

To retain and improve their utility as decision aids, health economic models require updating and refinement as new evidence emerges and decision contexts change [@Jenkins2021]. The significant deficits in our understanding of the systems in which mental disorders emerge and are treated [@Fried2020] suggest that there is ample scope for mental health systems models to progressively improve the validity over time. Prospective work could address the weak theoretical underpinnings for understanding complex mental health systems [@RN2111]. For example, it remains unclear why increased investments in mental health care have yet to discernibly reduce the prevalence and burden of mental disorders[@RN26].  The literature about how the requirements, characteristics and performance of mental health services are shaped by spatiotemporal context needs to be further developed [@RN42]. There is also a need for better evidence to identify the social determinants of mental disorders most amenable to preventative interventions, and for which population sub-groups such interventions would be most effective [@RN43].

The development, validation and updating of more complex mental health economic models implemented over longer time frames may be too onerous a burden for a single modelling team. Developing networks of modellers working on common health conditions has been recommended as a strategy for improving model validity [@Sampson2019] and some of us are part of a nascent initiate of this type in mental health [@nmhsmn2022]. Furthermore, collaborations across multiple modelling teams that include the ability to re-use and extend each others work can make complex modelling projects more tractable [@Arnold2010]. <!-- Add ref to NMHSMN in Australia --> Similarly, developing partnerships between modellers and decision-makers across the life-cycle of a modelling project can help ensure models are appropriately conceptualised and implemented and improve their practical utility as decision aids [@Zabell2021; @SQUIRES2016588].

@@

Our initial work on **readyforwhatsnext** is focused on Victoria, Australia but the framework we are using to develop it is designed to facilitate extension by ourselves and others to different decision contexts. Progress is reported on a project website [@rfwn2022].

@@

## Framework standards
We have identified 20 standards that we believe are important for quality implementations of OSHEMs, each described under one of following six principles for making models **TIMELY**:

@@

Consistent use of meaningful naming conventions when authoring code is recommended [@Wilson_2017; @Alarid2019]. Code can be made easier to follow by using the practices of abstraction [@8717448], where only simple, high level commands are routinely exposed to reviewers, and polymorphsim [@7181447], where the same command (e.g. "simulate") can be reused to implement different algorithms of the same type. Programs to implement model analyses can be made comprehensible to even non-technical users though the use of literate programming techniques that use tools like RMarkdown [@xie2018r] to render documents that integrate computer code with plain English descriptions.

@@

Adopting semantic versioning [@semver20xx] conventions can be an efficient way to provide users of model code and data with information about the potential importance of an update. For R code, the usethis [@usethis2021] package can be used to partially automate version number increments using the convention Major.Minor.Patch.Development. Datasets stored on the Harvard Dataverse use the simpler Major.Minor convention. 
Continuous integration [@CI2017] tools can help verify that each code update passes multiple quality tests. OSHEMs developed in R can take advantage of templates provided by devtools [@devtools2021] and pkgdown [@pkgdown2022] to run continuous integration checks on GitHub. These tests can include those of units (do individual functions produce expected output?), documentation (does documentation render correctly?; can all example workflows be executed?) and installation (can the software be successfully deployed on multiple types of operating system?).

@@

### Modular Models
Modular health economic models link multiple self-contained components that can be independently reused and extended by other projects [@Trauer2017;@Urach2013].

Many types of mental health data are highly sensitive with strict confidentiality requirements. For this reason, not all data included in some mental health models can be made widely available for others to re-use. A modular approach that ensures that model code and data are decoupled (stored in different files) can help model developers to restrict access to confidential model data, while providing open access to all other model components.

@@

Writing code as collections of functions (short, self-contained and reusable algorithms that each perform a discrete task) is recommended as good practice for scientific computing [@Wilson_2017]. When distributed as libraries (for example, as R packages), functions have the potential to be widely re-used, often in contexts very different than those they were originally developed for. A special type of function, called a method, can only be applied to a pre-defined class of data structure. Due to the coding concept of inheritance [@8717448], the more restricted nature of methods can be used to enhance the re-usability of model code in different decision contexts [@ready4oop2022]. For example, when generalising a model developed for the Australian context to a UK context, one could create a class that initially inherits all of the methods defined for the Australian model and then write new or replacement methods as needed for the UK model.

Whatever type of functions are written for a modelling project, it is good practice to make available test or toy data to demonstrate their use [@Wilson_2017].

@@

To make model code and data widely re-usable by others, it is important to provide users with appropriate and explicit permissions. In the context of open source models, there are two broad categories of licensing options. Some guidance strongly recommends the use of permissive licensing [@Wilson_2017] that provides users with great flexibility as to the purposes (including commercial) for which the content could be re-used. An alternative approach is to use copyleft licenses [@copyleft2022] that can require content users to distribute any derivative works they create under similar open source arrangements.

@@

Statistical models are a common output of health economic evaluations, but they are often not reported in a format that enables others to confidently and reliably re-use them [@Kearns2013]. Open source approaches can help address this by disseminating code artefacts that enable easy and appropriate use of a statistical model to make predictions with new data. However, great care must be exercised when doing so if models are derived from data on human subjects. In the R language, model objects by default typically contain a copy of the source dataset. Such dataset copies must therefore be replaced (for example, with fake data) and the amended artefact's predictive performance then retested before any public release.

@@

The six R packages, their primary focus, the TIMELY standards they support and the third-party packages they depend on are summarised in Table \@ref(tab:cpkgs). When used in conjunction with framework repositories, the six packages extending existing R packages to provide strong support for implementing 16 of the TIMELY standards. However, the software only weakly supports implementing the standards relating to disseminating statistical models (Y1) and user-interface development (Y2) and does not yet provide any workflow tools to help implement the standards for code coverage (T5) and deprecation conventions (I4). Standards not supported or weakly supported by our software can be met with existing developer tools in R and we plan to progressively integrate these third-party tools with our own in future releases of our software. Another future priority is to submit production versions of each R package for review by and archiving on the Comprehensive R Archive Network (CRAN) [@CRAN2022].

>>>>

Mental disorders impose high health, social and economic burdens worldwide [@RN8;@GBD2019]. Much of this burden is potentially avertable [@RN25], but poorly financed and organised mental health systems are ill-equipped for this challenge[@RN22;@RN23]. The large and widespread additional mental health burdens recently observed during the COVID-19 pandemic[@20211700] and predicted as a potential future consequence of climate change [@page_howard_2010], highlight the need to improve the resilience and adaptability of these systems. To help stem growing demand for mental health services, policymakers have also been encouraged to place greater emphasis on tackling the social determinants of mental disorders [@RN11].

<!-- Computational modelling could play an important role in developing policies to improve population mental health but this may require significant changes in the way mental health modelling projects are funded, conceptualised and implemented.  -->
Open source frameworks have been recommended to develop the mental health modelling field [@RN73] but, as with health economics more generally, OSHEMs remain rare. Currently there is only one mental health related model (in Alcohol Use Disorder [@Basu2018]) that is indexed in the Open Source Models Clearinghouse [@OSMC_20xx;@Emerson2019]. An open source Major Depressive Disorder reference model for the United States [@IVIMDD2022] is also in development <!-- as part of the Open Value Initiative [@Jansen2019].  -->Greater use of open source approaches could help improve the scope, validity and usefulness of mental health economic models.

Major mental health reform programs, <!-- such as those currently being implemented in Victoria, Australia [@VicRC2021],  -->can involve the identification, prioritisation, sequencing, targeting and monitoring of multiple interdependent initiatives. Single purpose models that assume static systems may be inadequate for the decision support needs of policymakers and service planners [@PC2020]. Mental health simulation studies rarely model complex dynamic systems [@RN73] with economic studies predominantly examining the affordability and value for money of individual mental health programs [@RN34]. Systems modelling approaches, that have been recommended for public health economics more generally [@SQUIRES2016588] and for mental health specifically [@Occhipinti2021], could provide greater insight about inter-dependencies between candidate policies and the dynamic nature of the mental health systems planning context. Dynamic systems methods might provide the foundations for developing reference models [@Afzali2013] of mental health systems that are intended for multiple-applications and re-use by different modelling teams. However, more complex models may be more prone to error [@Saltelli2019] and models designed for multiple purposes require greater investments in model transparency and validation [@Eddy2012; @Feenstra2022].

To retain and improve their utility as decision aids, health economic models require updating and refinement as new evidence emerges and decision contexts change [@Jenkins2021]. The significant deficits in our understanding of the systems in which mental disorders emerge and are treated [@Fried2020] suggest that there is ample scope for mental health systems models to progressively improve their validity over time. Prospective work could address the weak theoretical underpinnings for understanding complex mental health systems [@RN2111]. For example, it remains unclear why increased investments in mental health care have yet to discernibly reduce the prevalence and burden of mental disorders[@RN26].  The  literature, and evidence base, regarding how the requirements, characteristics and performance of mental health services are shaped by spatiotemporal context needs to be further developed [@RN42]. There is also a need for better evidence to identify the social determinants of mental disorders most amenable to preventative interventions, and for which population sub-groups such interventions would be most effective [@RN43].

The development, validation and updating of more complex mental health economic models implemented over longer time frames may be too onerous a burden for a single modelling team. Developing networks of modellers working on common health conditions has been recommended as a strategy for improving model validity [@Sampson2019] and some of us are part of a nascent initiate of this type in mental health [@nmhsmn2022]. Furthermore, collaborations across multiple modelling teams that include the ability to re-use and extend one another's work can make complex modelling projects more tractable [@Arnold2010]. <!-- Add ref to NMHSMN in Australia --> Similarly, developing partnerships between modellers and decision-makers across the life-cycle of a modelling project can help ensure models are appropriately conceptualised, implemented and have practical utility as decision aids [@Zabell2021; @SQUIRES2016588].

@@

Our initial work on **readyforwhatsnext** is focused on Victoria, Australia but the framework we are using to develop it is designed to facilitate extension by ourselves and others to different jurisdictional decision contexts. Progress is reported on a project website [@rfwn2022].

@@

## Framework standards
Based on existing open science guidance and our own experience, we have identified 20 standards that are important for quality implementations of OSHEMs, each described under one of following six principles for making models **TIMELY**:

@@

Consistent use of meaningful naming conventions when authoring code is recommended [@Wilson_2017; @Alarid2019]. Code can be made easier to follow by using the practices of abstraction [@8717448], where only simple, high level commands are routinely exposed to reviewers, and polymorphism [@7181447], where the same command (e.g. "simulate") can be reused to implement different algorithms of the same type. Programs to implement model analyses can be made comprehensible to even non-technical users through the use of literate programming techniques that use tools like RMarkdown [@xie2018r] to render documents that integrate computer code with plain English descriptions.

@@

Adopting semantic versioning [@semver20xx] conventions can be an efficient way to provide users of model code and data with information about the potential importance of an update. For R code, the usethis [@usethis2021] package can be used to partially automate version number increments using the convention Major.Minor.Patch.Development (e.g. ExampleSoftware v1.1.1.1). Datasets stored on the Harvard Dataverse use the simpler Major.Minor convention.

Continuous integration [@CI2017] tools can help verify that each code update passes multiple quality tests. OSHEMs developed in R can take advantage of templates provided by devtools [@devtools2021] and pkgdown [@pkgdown2022] to run continuous integration checks on GitHub. These tests can include those of units (do individual functions produce expected output?), documentation (does documentation render correctly?; can all example workflows be executed?) and installation (can the software be successfully deployed on multiple types of operating system?).

@@

### Modular Models
Modular health economic models link multiple self-contained components that can be independently reused and extended by other projects [@Trauer2017;@Urach2013]. Many types of mental health data are highly sensitive with strict confidentiality requirements. For this reason, not all data included in some mental health models can be made widely available for others to re-use. A modular approach that ensures that model code and data are decoupled (stored in different files) can help model developers to restrict access to confidential model data, while providing open access to all other model components.

@@

Writing code as collections of functions (short, self-contained and reusable algorithms that each perform a discrete task) is recommended as good practice for scientific computing [@Wilson_2017]. When distributed as libraries (for example, as R packages), functions have the potential to be widely re-used, often in contexts very different than those they were originally developed for. A special type of function, called a method, can only be applied to a pre-defined class of data structure. Due to the coding concept of inheritance [@8717448], the more restricted nature of methods can be used to enhance the re-usability of model code in different decision contexts [@ready4oop2022]. For example, when generalising a model developed for the Australian context to a UK context, one could create a class that initially inherits all of the methods defined for the Australian model and then write new or replacement methods as needed for the UK model.Whatever type of functions are written for a modelling project, it is good practice to make available test or toy data to demonstrate their use [@Wilson_2017].

@@

To make model code and data widely re-usable by others, it is important to provide users with appropriate and explicit permissions. In the context of open source models, there are two broad categories of licensing options. Some guidance strongly recommends the use of permissive licensing [@Wilson_2017] that provides users with great flexibility as to the purposes (including commercial) for which the content could be re-used. An alternative approach is to use copyleft licenses [@copyleft2022] that can require content users to distribute any derivative works they create under similar open source arrangements.The most suitable approach may differ for code and data.

@@

Statistical models are a common output of health economic evaluations, but they are often not reported in a format that enables others to confidently and reliably re-use them [@Kearns2013]. Open source approaches can help address this by disseminating code artefacts that enable easy and appropriate use of a statistical model to make predictions with new data. However, great care must be exercised when doing so if models are derived from data on human subjects. In the R language, model objects by default typically contain a copy of the source dataset. Such dataset copies must therefore be replaced (for example, with synthetic data) and the amended artefact's predictive performance then retested before any public release.

@@

The six R packages, their primary focus, the TIMELY standards they support and the third-party packages they depend on are summarised in Table \@ref(tab:cpkgs). When used in conjunction with framework repositories, the six packages extend existing R packages provide strong support for implementing 16 of the TIMELY standards. However, the software only weakly supports implementing the standards relating to disseminating statistical models (Y1) and user-interface development (Y2) and does not yet provide any workflow tools to help implement the standards for code coverage (T5) and deprecation conventions (I4). Standards not supported or weakly supported by our software can be met with existing developer tools in R and we plan to progressively integrate these third-party tools with our own in future releases of our software. Another future priority is to submit production versions of each R package for review by and archiving on the Comprehensive R Archive Network (CRAN) [@CRAN2022].
